---
title: "Moon Days Data Analysis"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
```

I've taken the Maha Nikaya icalendar and for each event written the date and
event summary to a CSV file. 

## Import the CSV
```{r import-csv }
events <- read_csv("events.csv", col_names = c("summary_text", "date"))
```

## Examine the data

To start with, let's see what the raw data looks like.
```{r examine }
events %>% head(10)
```

## What is the range of dates?
```{r date-range }
date_range <- events %>% 
  summarise(
    start = min(date),
    end = max(date)
  )
```

The event list covers the dates from `r date_range$start` to `r date_range$end`.

## What are the distinct types of event

```{r event-types }
events %>% count(summary_text, sort = TRUE)
```

There are the waxing & waning 1/2 moon days, holidays, rains retreat dates and the uposathas which include:

- Full/New
- 14/15 day
- Season in Pali
- Week in season & number of weeks in season.

Some parsing of the summaries will be in order.

## Multiple summaries on same date

In some cases there are more than one event. What does that look like?
```{r multiple summaries }
multiple_summaries <- events %>% 
  group_by(date) %>%
  summarise(
    n = n()
  ) %>%
  filter(n > 1)

multi_events <- events %>% 
  semi_join(multiple_summaries, by = "date")

multi_events %>%
  mutate(date = stamp("1 Jan 2022", quiet = TRUE)(date)) %>%
  select(date, summary_text) %>%
  head(15) %>%
  knitr::kable(col.names = c("Date", "Summary Text"))
```
### Intepretation

It looks like there dates with 2 summaries where the uposatha is also a
special day. There are some with 3 summaries.

## Dates with 3 summaries
```{r 3 summaries}
multiple_summaries <- events %>% 
  group_by(date) %>%
  summarise(
    n = n()
  ) %>%
  filter(n == 3)

three_summary_events <- events %>% 
  semi_join(multiple_summaries, by = "date")

three_summary_events %>%
  mutate(date = format(date, "%d %h %Y")) %>%
  select(date, summary_text) %>%
  knitr::kable(col.names = c("Date", "Summary Text"))
```
### Interpretation

Here we have three bits of information - the uposatha, the pavarana and the
fact that it is the last day of the vassa. 

To build our python module, we can make some assumptions about the structure of
the data, in particular the use of the summary field. It has been used in a
very consistent way.

## Waxing & waning moons

Are these events always by themselves?
```{r waxing-waning }
waning_waxing <- events %>%
  filter(summary_text %in% c("Waxing Moon", "Waning Moon"))

events %>% 
  semi_join(waning_waxing, by = "date") %>%
  group_by(date) %>%
  summarise(
    n = n()
  ) %>%
  filter(n > 1)
```
Yes, that's right.

## Special days

Do special days always fall on an uposatha?
```{r ordinary}
ordinary <- events %>%
  filter(str_detect(summary_text, "Moon"))

ordinary %>% count(summary_text)
```
### Get the special days
```{r special}
special <- events %>%
  anti_join(ordinary, by = "summary_text") %>%
  filter(!(str_detect(summary_text, "Vassa")))

special %>% count(summary_text)
```
### What events are concurent with the special days?
```{r concurent-with-special }
concurent_with_special <- events %>% 
  semi_join(special, by = "date") %>%
  filter(!str_detect(summary_text, "Vassa"))

concurent_with_special %>% count(summary_text)
```
### Interpretation

All special days fall on the full moon day.

## Rains beginning and end days
```{r vassa-days }
vassa_days <- events %>% filter(str_detect(summary_text, "Vassa"))

same_day_as_vassa_days <- events %>% semi_join(vassa_days, by = "date")

same_day_as_vassa_days %>% count(summary_text)
```
So the end of rains includes three events. Let's just check that the first day
of vassa stands alone, is not accompanied by a moon day or special day.
```{r first-day-of-vassa}
first_day_of_vassa <- events %>% filter(summary_text == "First day of Vassa")
same_as_first_day_of_vassa <- events %>% semi_join(first_day_of_vassa, by = "date")

same_as_first_day_of_vassa %>% count(summary_text)
```

It appears that the last day of the vassa is not a moon day or a special day.

## What assumptions can we make?

**Events occurring by themselves**

- Waxing moon
- Waning moon
- New moon
- Full moon
- First day of vassa

**More than one event on same date**

- Full moon + special day
- Full moon + special day (pavarana) + last day of vassa

## Are dates sequential?

If they are, we can keep an event object and add the special/vassa days to it
as we walk through the icalendar. As long as the date is the same we just add
it to the most recent event. If the dates are not sequential, we need to check whether we already
have an event at that date before adding the extra information. Also, if we want to add extra 
information to a waxing/waning moon event (season, week in season) we (might?) need to 
search the list of events to find the previous date. Starting with a sequential 
list we can just keep track of the previous event.

**OK, I'm  a bit rusty when it comes to algorithms.**

```{r are-dates-sequential }
events %>%
  mutate(
    next_out_of_sequence = (lead(date) < date)
  ) %>%
  count(next_out_of_sequence)
```
That one NA is the last in the sequence and has no next event. 

**Assumption: dates are sequential**
