---
title: "Reviewing assumptions"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
```

## Get the data
```{r import }
events <- read_csv("events.csv", col_names = c("summary_text", "date")) %>%
  select(date, summary_text) 
```

In the first analysis we arrived at a number of assumptions about the icalendar
file from which we draw our data. Here I attempt to confirm each one.

## Dates are in order
```{r tag-sequential }
sequential <- events %>%
  mutate(
    is_in_sequence = (lag(date) <= date)
  )

sequential %>% count(is_in_sequence)
```

All dates are sequential, keeping in mind `lag(date)` is `NA` for the first row. 

## Some types of event stand alone

Waxing moon, waning moon and first day of vassa never occur with another summary. 
We can count the summary text entries for the dates these occur to confirm this

**Oops, forgot new moons. Have to sort that out.**
```{r solitary }
solitary <- events %>% 
  filter(summary_text %in% c("Waning Moon", "Waxing Moon", "First day of Vassa"))

events %>% 
  semi_join(solitary, by = "date") %>%
  count(date) %>%
  filter(n > 1)
```

## Create a tidy view

Here we make a table to work with.

```{r event_types }
summary_types <- events %>%
  mutate(
    summary_type = if_else(str_detect(summary_text, "Full Moon"), "Full", "None"),
    summary_type = if_else(str_detect(summary_text, "New Moon"), "New", summary_type),
    summary_type = if_else(str_detect(summary_text, "Waxing Moon"), "Waxing", summary_type),
    summary_type = if_else(str_detect(summary_text, "Waning Moon"), "Waning", summary_type),
    summary_type = if_else(str_detect(summary_text, "First day of Vassa"), 
                           "First day of Vassa", summary_type),
    summary_type = if_else(str_detect(summary_text, "Last day of Vassa"), 
                           "Last day of Vassa", summary_type),
    summary_type = if_else(summary_text == "Māgha Pūjā", "Special", summary_type),
    summary_type = if_else(summary_text == "Visākha Pūjā", "Special", summary_type),
    summary_type = if_else(summary_text == "Āsāḷha Pūjā", "Special", summary_type),
    summary_type = if_else(summary_text == "Pavāraṇā Day", "Special", summary_type),
  )

summary_types %>% count(summary_type)
```
Now we get the events per date.
```{r events-per-date }
summaries_per_date <- summary_types %>%
  group_by(date) %>%
  summarise(
    summary_type,
    summary_text_count = n()
  ) 

summaries_per_date
```

```{r varying-number-of-summaries }
variation_by_type <- summaries_per_date %>%
  group_by(summary_type) %>%
  count(summary_text_count)

variation_by_type
```

## Finally an easy to read table
```{r pivot-wider }
wider_variation_by_type <- variation_by_type %>%
  pivot_wider(names_from = summary_text_count, values_from = n)

wider_variation_by_type
```